---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# rTPC

The aim of rTPC is to help fit thermal performance in R. These functions act as wrappers that aid in the fitting of common functions for fitting thermal performance curves.

## Installation

```{r install, eval = FALSE}
# install package from GitHub
devtools::install_github("padpadpadpad/rTPC")
```

To show each function, we shall fit the models using nls.multstart. To do we shall install other packages that will help us.

```{r install2, eval = FALSE}
# install nls.multstart
install.packages('nls.multstart')
install.packages('purrr')
install.packages('dplyr')
install.packages('tidyr')
install.packages('ggplot2')
install.packages('nls.multstart')
install.packages('broom')
```

We can now load in the data and the packages to fit the TPCs

```{r quicklook, out.width = '50%', fig.height = 4, fig.width = 6, fig.align='center', message=FALSE}
# load in packages
library(purrr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(nls.multstart)
library(broom)
library(rTPC)

# write function to label ggplot2 panels
label_facets <- function(string){
  len <- length(string)
  string = paste('(', letters[1:len], ') ', string, sep = '')
  return(string)
}

# write function to convert label text size to points
pts <- function(x){
  as.numeric(grid::convertX(grid::unit(x, 'points'), 'mm'))
}

# load in data
data("Chlorella_TRC")

# change rate to be non-log transformed
d <- mutate(Chlorella_TRC, rate = exp(ln.rate))

# filter for just the first curve
d_1 <- filter(d, curve_id == 1)

# plot
ggplot(d_1, aes(temp, rate)) +
  geom_point() +
  theme_bw(base_size = 16)
```

We can now run nls.multstart for each function

```{r fit_models}
# run in purrr - going to be a huge long command this one
d_models <- group_by(d_1, curve_id, growth.temp, process, flux) %>%
  nest() %>%
  mutate(., lactin2 = map(data, ~nls_multstart(rate ~ lactin2_1995(temp = temp, p, c, tmax, tref),
                       data = .x,
                       iter = 500,
                       start_lower = c(p = 0, c = -2, tmax = 35, tref = 0),
                       start_upper = c(p = 3, c = 0, tmax = 55, tref = 15),
                       supp_errors = 'Y')),
            sharpeschoolhigh = map(data, ~nls_multstart(rate ~ sharpeschoolhigh_1981(temp_k = K, r_tref, e, eh, th, tref = 15),
                                                     data = .x,
                                                     iter = 500,
                                                     start_lower = c(r_tref = 0.01, e = 0, eh = 0, th = 270),
                                                     start_upper = c(r_tref = 2, e = 3, eh = 10, th = 330),
                                                     supp_errors = 'Y')),
            thomas = map(data, ~nls_multstart(rate ~ thomas_2012(temp = temp, a, b, c, topt),
                                           data = .x,
                                           iter = 500,
                                           start_lower = c(a = -10, b = -10, c = -10, topt = 0),
                                           start_upper = c(a = 10, b = 10, c = 10, topt = 40),
                                           supp_errors = 'Y')),
            briere2 = map(data, ~nls_multstart(rate ~ briere2_1999(temp = temp, tmin, tmax, a, b),
                                           data = .x,
                                           iter = 500,
                                           start_lower = c(tmin = 0, tmax = 20, a = -10, b = -10),
                                           start_upper = c(tmin = 20, tmax = 50, a = 10, b = 10),
                                           supp_errors = 'Y',
                                           lower = c(tmin = -10, tmax = 20, a = -10, b = -10),
                                           upper = c(tmin = 20, tmax = 80, a = 10, b = 10))),
            spain = map(data, ~nls_multstart(rate ~ spain_1982(temp = temp, a, b, c, r0),
                                           data = .x,
                                           iter = 500,
                                           start_lower = c(a = -1, b = -1, c = -1, r0 = -1),
                                           start_upper = c(a = 1, b = 1, c = 1, r0 = 1),
                                           supp_errors = 'Y')),
            ratkowsky = map(data, ~nls_multstart(rate ~ ratkowsky_1983(temp = temp, tmin, tmax, a, b),
                                           data = .x,
                                           iter = 500,
                                           start_lower = c(tmin = 0, tmax = 20, a = -10, b = -10),
                                           start_upper = c(tmin = 20, tmax = 50, a = 10, b = 10),
                                           supp_errors = 'Y')),
            boatman = map(data, ~nls_multstart(rate ~ boatman_2017(temp = temp, rmax, tmin, tmax, a, b),
                                           data = .x,
                                           iter = 500,
                                           start_lower = c(rmax = 0, tmin = 0, tmax = 20, a = -1, b = -1),
                                           start_upper = c(rmax = 2, tmin = 20, tmax = 50, a = 1, b = 1),
                                           supp_errors = 'Y')))

```

We can now make predictions of each model and plot them. In addition, we can estimate optimum temperature for each model using `get_topt()`.

```{r plot_predictions, fig.height = 10, fig.width = 10}
# stack models
d_stack <- gather(d_models, 'model', 'output', 6:ncol(d_models))

# preds
newdata <- tibble(temp = seq(min(d_1$temp), max(d_1$temp), length.out = 100),
                  K = seq(min(d_1$K), max(d_1$K), length.out = 100))
d_preds <- d_stack %>%
  unnest(., output %>% map(augment, newdata = newdata)) %>%
  mutate(., temp = ifelse(model == 'sharpeschoolhigh', K - 273.15, temp))

# estimate extra parameters
extra_params <- d_stack %>%
  mutate(., est = map(output, est_params)) %>%
  select(., -c(data, output)) %>%
  unnest(., est) %>%
  mutate(., topt = ifelse(topt > 200, topt - 273.15, topt),
         rmax = round(rmax, 2))

# plot
ggplot(d_preds, aes(temp, rate)) +
  geom_point(aes(temp, rate), d_1) +
  geom_text(aes(-Inf, Inf, label = paste('Topt =', topt, 'ºC\n', 'rmax = ', rmax)),  hjust = -0.1, vjust = 1.5, extra_params, size = pts(12)) +
  geom_line(aes(temp, .fitted, col = model)) +
  facet_wrap(~model, labeller = labeller(model = label_facets)) +
  theme_bw(base_size = 16) +
  theme(legend.position = 'none',
        strip.text = element_text(hjust = 0),
        strip.background = element_blank()) +
  xlab('Temperature (ºC)') +
  ylab('rate') +
  geom_hline(aes(yintercept = 0), linetype = 2)

```
