---
title: "Adding Models to rTPC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adding Models to rTPC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

So you want to contribute a model to rTPC? That's awesome!

Here we will go through all the major steps for creating and contributing a model to rTPC.

This vignette assumes that you have a working knowledge of basic git usage (either through the terminal, the RStudio interface, or another GUI), of R, and a functional knowledge of thermal performance curves.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
```{r runtimer, include=FALSE}
runstart <- lubridate::now()
```

---

## First steps

The first thing you want to do when you have a model that is potentially useful in rTPC is to _check whether it's already there_!

rTPC already contains many models designed on a variety of different data. It is quite likely at this point that there is a similar (or identical) model that already serves the same purpose. So why not save yourself the work if you can?

---

### Forking the repository

The easiest way to start working on rTPC is to **fork** the repo. [Github has some excellent instructions on doing this](https://docs.github.com/en/pull-requests/collaborating-with-pull-requests/working-with-forks/fork-a-repo).

Working on a fork of the repository will allow you to commit to rTPC as if it is your own project, and then contribute those commits back to the main repository through a **pull request**.

### Setting up your workspace and R installation

It is _strongly_ suggested that you install all packages from the "Suggests" section of rTPC (easily found in the DESCRIPTION file), alongside the packages `devtools` and `usethis` (see [here](https://r-pkgs.org/setup.html) for a deeper explanation of these packages).

Once you have done this, git clone from YOUR fork of rTPC into a folder of your choice, and then open up the `rTPC.Rproj` file if you are using RStudio.

---

## Adding a model

Adding a model to rTPC consists of 4 major steps:

1. Add the model code, starting values, and limits.
2. Add the test file.
3. Add the model to the rTPC database.
4. Clean up and check.

Models are generally named according to the scheme `authormodifier#_YYYY`. So a model by authors called King and Bishop published in 1992 would usually be codified as `kingbishop_1992`. If they further proposed a modified model, and a simplified version of each in the original paper, those would probably appear as 

- `kingbishopmodified_1992`
- `kingbishopsimplified_1992`
- `kingbishopmodifiedsimplified_1992`

If there are many authors, or another better description of the model then initialisms or other descriptors can be used, but these may be changed if the package maintainers deem it to be confusing.

---

### Model equation

For this example, let's implement the Eubank model from [The Significance and Thermodynamics of Fluctuating Versus Static Thermal Environments on _Heliothis zea_ Egg Development Rates](https://academic.oup.com/ee/article/2/4/491/2480483).

This model (eq4) is originally derived just for a bollworm, with some magic numbers:

$$R(T) = \frac{800}{(T-91)^2+432}$$

As a generalisation, we know from looking at the paper that $T_{opt} = 91$, and the numbers 800 and 432 are species-specific (and thus should be fit).

As such we can reformulate the equation as:

$$R(T) = \frac{a}{(T-T_{opt})^2+b}$$

Where $a$ and $b$ are arbitrary parameters for fitting, and $T_{opt}$ can be at least approximated from the data.

---

### Implementation

The model itself lives in a file named after the function, in the `R` directory. In our case that will be `eubank_1973.R`.

You can create one using the function `usethis::use_r("eubank_1973")`

Within this file are 4 functions:

- `eubank_1973()` - The model itself
- `eubank_1973.starting_vals()` - The starting values
- `eubank_1973.lower_lims()` - The lower limits
- `eubank_1973.upper_lims()` - The upper limits

It is _very important_ that the auxiliary functions are named in this manner, otherwise the unified starting value/limit dispatch will not function correctly.

---

#### Model function - `eubank_1973()`

First of all we must implement the actual model itself (i.e. the equation above, but as R code).

```{r echo=TRUE, eval=FALSE}
eubank_1973 <- function(temp, topt, a, b){
  est <- a / ((temp - topt)^2 + b)
  return(est)
}
```

A model function takes the parameters to fit the model as arguments, with temperature being the first by convention, and returns the estimates of the trait value at the given temperatures and parameter values.

---

#### Starting values - `eubank_1973.starting_vals()`

Deciding sensible starting values is truly a dark art. In many cases these are fairly arbitrary, though in certain cases an estimate can be obtained from the data.

Starting values functions are always passed a dataframe called `d`, which will contain only the temperature and trait values for the given curve:

```{r}
# A slightly compressed example of how d is generated
library(rTPC)
subs <- subset(chlorella_tpc, chlorella_tpc$curve_id == 1)
d <- data.frame(x=subs$temp, y=subs$rate, stringsAsFactors = FALSE)
d <- d[order(d$x),]
d
```

Every `model.starting_values()` function has the same approximate form, taking `d` as the only argument, and returning a named list of starting values. So in our case:

```{r, echo=TRUE, eval=FALSE}
eubank_1973.starting_vals <- function(d){
  # starting values go here
  return(c(topt=topt, a=a, b=b))
}
```


In the case of the eubank model, $a$ and $b$ are arbitrary. As such we can take fairly sensible values at approximately the right magnitude (say $a = 300$ and $b = 50$).

$T_{opt}$ on the other hand is a value that we can derive from our data, in this case it's the temperature at which $R$ is at its maximum.

So lets put that together in our starting_vals function

```{r}
eubank_1973.starting_vals <- function(d){
  rmax = max(d$y, na.rm = TRUE)  # Find max trait value
  topt = mean(d$x[d$y == rmax])  # Find T of rmax
  a = 300
  b = 50
  return(c(topt=topt, a=a, b=b))
}
```

To reiterate an earlier point, it is _essential_ that the starting values function is named appropriately, e.g. `eubank_1973.starting_vals()`.

The following _will not and can not work_:

- `eubank_1973.startingvals()`
- `eubank_1973.start_vals()`
- `eubank_1973_starting_vals()`
- `.starting_vals()`
- `model.starting_vals()`

---

#### Limits - `eubank_1973.lower_lims()` & `eubank_1973.upper_lims()`

NLLS fitting often fails when unconstrained (as it may take one parameter to infinity and ignore all others). Even specifying absurdly large or infinite limits can make the optimiser behave better.

As such, we must provide realistic bounds on the upper and lower limits of each parameter.

These take the same functional form as the `starting_vals` function, requiring one argument (`d`) and returning a named list.

Generally speaking it is best to keep limits fairly wide unless there is a good biological reason to do otherwise.

For instance at the most $T_{opt}$ is never going to be above about 150°C, and likely will be a lot lower. And at its lower bound it is unlikely for the optimal trait value to occur at temperatures below 0°C. That gives us some fairly reasonable bounds for the values of $T_{opt}$.

$a$ and $b$, however, are arbitrary values. Just by looking at the functional form of the model we can see that if $a$ goes to or below 0, or $b = 0-(T-T_{opt})$ then we will end up with fairly nonsensical answers (either negative trait values or undefined results), so we can bound those both on the low end at 0 to be safe. 

In terms of upper limits, we have no real idea of how large those values may go (as they will be proportional to the units of the trait), so we can't set any reasonable bounds aside from `Inf`.

To put that together, here are the lower and upper limit functions for `eubank_1976`:

```{r}
eubank_1973.lower_lims <- function(d){
  topt = 0
  a = 0
  b = 0
  return(c(topt=topt, a=a, b=b))
}

eubank_1973.upper_lims <- function(d){
  topt = 150
  a = Inf
  b = Inf
  return(c(topt=topt, a=a, b=b))
}
```

---

#### Documentation

rTPC leverages `roxygen2` for documentation. Each file contains documentation definitions at the **top of the file** in special comments starting `#'`.

The best way to implement these is to copy from another model file (such as `atkin_2005`) and modify it to suit your own needs.

Things to modify:

- The first line description
- parameters
- author
- reference
- equation
- note (if poorly fitting)
- examples
- export directive (this _must_ match the name of the model! The starting value and limit functions do not need exporting)

---

### Updating the model database

Once you have created a model file, you need to add this model to rTPC's database of model names, so it knows that the model is available.

This simply requires modifying the `mod_names` vector within the `get_model_names.R` file to include your new model. Make sure that the model name is exactly the same as that in your original file, and that it is enclosed in quotes!

---

### Testing model implementation

Now that the model is set up, the next port of call is to simply run the integration test.

This is located in the `tests/testthat/test-startingvalues.R`

It should provide a result somewhat like.
```
[ FAIL 0 | WARN 0 | SKIP 0 | PASS 3 ]
```

If the test failed, the output is instead more like this:

```
[ FAIL 3 | WARN 0 | SKIP 0 | PASS 0 ]

── Failure (test-startingvalues.R:42:3): All models can generate starting values ──
length(mod_names) not equal to `starting_count`.
1/1 mismatches
[1] 48 - 47 == 1
No starting values for:
 ebank_1973

── Failure (test-startingvalues.R:46:3): All models can generate lower limits ──
length(mod_names) not equal to `lower_count`.
1/1 mismatches
[1] 48 - 47 == 1
No lower limits for:
 ebank_1973

── Failure (test-startingvalues.R:50:3): All models can generate upper limits ──
length(mod_names) not equal to `upper_count`.
1/1 mismatches
[1] 48 - 47 == 1
No upper limits for:
 ebank_1973
[ FAIL 3 | WARN 0 | SKIP 0 | PASS 0 ]
```

In this case we can see that I simply misspelled the name of the model in the `get_model_names.R` file, so need to change that.

If only one test of the three failed, that is likely an issue with the names of the functions in the actual model file.

---

### Individual model test

Each model requires a model test, to make sure that changes in the future do not cause major fitting issues. These test files live in the `tests/testthat` directory.

Creating a new one in RStudio is easy! Just make sure you have your model file in the script window, then run the command `usethis::use_test()`.

This will create a new test file, in our case `tests/testthat/test-eubank_1973.R`

Tests are mostly boilerplate, however there are a few things that do need to be changed, so again you can copy most of the code from another test file, such as `test-atkin_2005.R`.

The main items to be changed are:

- The model name throughout (5 occurrences)
- The model arguments within `nls_multstart`
- The `iter` argument to `nls_multstart` (this should be something like `c(3,3,3,3)` for a model with 4 free parameters)

So the final test file will look something like this:

```{r echo=TRUE, eval=FALSE}
# do not run the test on CRAN as they take too long
testthat::skip_on_cran()

# method: fit model and get predictions. Check these against others.

# load in ggplot
library(ggplot2)

# subset for the first TPC curve
data('chlorella_tpc')
d <- subset(chlorella_tpc, curve_id == 1)

# get start values and fit model
start_vals <- get_start_vals(d$temp, d$rate, model_name = 'eubank_1973')

# fit model
mod <- nls.multstart::nls_multstart(rate~eubank_1973(temp = temp, tops, a, b),
                                    data = d,
                                    iter = c(3,3,3),
                                    start_lower = start_vals - 10,
                                    start_upper = start_vals + 10,
                                    lower = get_lower_lims(d$temp, d$rate, model_name = 'eubank_1973'),
                                    upper = get_upper_lims(d$temp, d$rate, model_name = 'eubank_1973'),
                                    supp_errors = 'Y',
                                    convergence_count = FALSE)

# get predictions
preds <- broom::augment(mod)
# dput(round(preds$.fitted, 1))

# plot
ggplot(preds) +
  geom_point(aes(temp, rate)) +
  geom_line(aes(temp, .fitted)) +
  theme_bw()

# run test
testthat::test_that("eubank_1973 function works", {
  testthat::expect_equal(
    round(preds$.fitted, 1),
    c(0.2, 0.2, 0.3, 0.4, 0.6, 0.9, 1.3, 1.6, 1.4, 1, 0.7, 0.4))
})
```

---

#### Getting results values

The second-from-final line in the test file defines the expected return from a model run. It is not really feasible to know this ahead of time, but luckily the code here contains an extra function that makes retrieving the expected value easy.

Before running the test for the first time, uncomment the line `dput(round(preds$.fitted, 1))`.

Now when you run the test with `testthat::test_file("tests/testthat/test-eubank_1973.R")` then the test will very likely fail, but it will also output a vector to the console.

```
[ FAIL 0 | WARN 0 | SKIP 0 | PASS 0 ]c(0.2, 0.2, 0.3, 0.4, 0.6, 0.9, 1.3, 1.6, 1.4, 1, 0.7, 0.4)
[ FAIL 1 | WARN 0 | SKIP 0 | PASS 0 ]

── Failure (test-eubank_1973.R:39:3): eubank_1973 function works ───────────────
round(preds$.fitted, 1) not equal to c(0.2, 0.3, 0.5, 0.6, 0.8, 1, 1.1, 1.2, 1.3, 1.2, 1, 0.1).
11/12 mismatches (average diff: 0.209)
[2]  0.2 - 0.3 == -0.1
[3]  0.3 - 0.5 == -0.2
[4]  0.4 - 0.6 == -0.2
[5]  0.6 - 0.8 == -0.2
[6]  0.9 - 1.0 == -0.1
[7]  1.3 - 1.1 ==  0.2
[8]  1.6 - 1.2 ==  0.4
[9]  1.4 - 1.3 ==  0.1
[10] 1.0 - 1.2 == -0.2
...
[ FAIL 1 | WARN 0 | SKIP 0 | PASS 0 ]
```

You can paste this vector (in this case `c(0.2, 0.2, 0.3, 0.4, 0.6, 0.9, 1.3, 1.6, 1.4, 1, 0.7, 0.4)`) onto the second from last line (remembering the extra closing parenthesis), then save and run again.

On this second run the test should execute properly.

```
[ FAIL 0 | WARN 0 | SKIP 0 | PASS 1 ]
```

You can then proceed with commenting back out the `dput()` line.

---

## Testing

At this point it is worth doing a bit of testing of your new function.

Load your new version of the package with `devtools::load_all(".")` and then in a new R (or Rmarkdown) file try running some tests with your new function.

If you are having trouble with fitting, modify your starting values or limits and try again until you get fairly good fits.

You may also have to play with the `iter` argument to `nls_multstart`.

Once you are happy with everything it's time to run R CMD CHECK (for example using `devtools::check()`).

Look to make sure that there are no warnings or errors.

```
── R CMD check results ──
Duration: 4m 55.6s

0 errors ✔ | 0 warnings ✔ | 0 notes ✔

R CMD check succeeded
```

---

## Documenting your model

Documentation needs to be generated in order to be available for your own model.

To do this, simply run the command `devtools::document()`.

Once this has completed you can view your documentation internally using `?modelname_yyyy`.

---

## Commiting your changes back upstream

Now you have fully implemented your model it is time to hand it back to the rTPC team for integration.

The simplest way to do this is to commit your changes (including the .Rd documentation file), push those to your own forked repo, then open a pull request (a prompt should appear on your repo on the github website).

Please make sure you include a sensible title such as "Add eubank_1976 model". 

Include in the description any notes about the model itself, decisions you may have had to make when creating it, or problems and difficulties when fitting the model.

See [PR #58 ](https://github.com/padpadpadpad/rTPC/pull/58) for an example of a good pull request.

---

## Done!

Now you are done and can go and have a well-earned cup of tea! You deserve it!

---

## Auxiliary model function dispatch

_A note on **how** rTPC dispatches auxiliary model functions_

Under the hood, rTPC relies on the `do.call()` function to locate and run the appropriate model functions.

A good example of this is when getting starting values.

The `get_start_vals()` function in its entirety is replicated here:

```{r eval=FALSE, echo=TRUE}
get_start_vals <- function(x, y, model_name) {

  mod_names <- get_model_names(returnall = TRUE)
  model_name <- tryCatch(rlang::arg_match(model_name, mod_names), error = function(e){
    cli::cli_abort(c("x"="Supplied {.arg model_name} ({.val {model_name}}) is not an available model in rTPC.",
                     "!"="Please check the spelling of {.arg model_name}.",
                     " "="(run {.fn rTPC::get_model_names} to see all valid names.)",
                     ""), call=rlang::caller_env(n=4))
  })

  # make data frame
  d <- data.frame(x, y, stringsAsFactors = FALSE)
  d <- d[order(d$x),]

  start_vals <- tryCatch(do.call(paste0(model_name, ".starting_vals"), list(d=d)),
                         error = function(e){NULL})

  return(start_vals)
}
```

Here the model name is checked against the list of valid models, and a source data frame (`d`) is created.

Finally a model function name is created by pasting together the name of the model along with `".starting_vals"`:

```{r}
model_name <- "eubank_1976"
paste0(model_name, ".starting_vals")
```

This function name is then executed using `do.call()` along with a list of arguments (in this case simply `d`), and the value is returned.

If a model function is not present or is spelled incorrectly, `do.call()` will throw an error which the `tryCatch` will convert into a `NULL` output.

In the case of an entirely missing model, `get_start_vals()` errors out early with a more specific error message.

```{r tot_time, include=FALSE}
tot_time <- lubridate::as.duration(lubridate::now() - runstart)
```

<span style="opacity: 0.1;font-size: small;">Built in `r tot_time`s</span>
