---
title: "Parallelising rTPC"
author: "Daniel Padfield"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Parallelising rTPC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#### A brief example of how to parallelise fitting many models to many TPCs using rTPC, nls.multstart the tidyverse, and mirai.

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  #tidy.opts=list(width.cutoff=60),
  #tidy=TRUE,
  fig.align = 'center'
)
```

***

## Things to consider

- **quickfit_tpc_multi()** makes assumptions about start values and currently does not support functions where you have to specify a reference temperature (e.g. **sharpe_schoolfield_1981()**). This is **by design** because you should think hard about the fitting process and writing your own functions will help you understand the process.
- We recommend writing your own functions for fitting models to pass to **purrr**.

```{r setup, message=FALSE}
# load in packages
library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)
library(mirai)

# load in data
data('chlorella_tpc')
d <- chlorella_tpc
head(d)
```

In our previous vignettes, we have shown how to fit many models to many curves using **rTPC**, **nls.multstart()**, and the **tidyverse**. However, as your datasets get larger (e.g. tens, hundreds, or thousands of TPCs) this could take a long time to run.

Excitingly, [**purrr**](https://www.tidyverse.org/blog/2025/07/purrr-1-1-0-parallel/) now supports parallelisation when combined with [**mirai**](https://mirai.r-lib.org/). This is the reason why we recommend using a single **purrr::map()** call that encompasses all models, as it integrates much better with **mirai** and allows for parallelisation.

Even more excitingly, we have functions in **rTPC** that allow you to easily fit one - **quickfit_tpc()** - or multiple - **quickfit_multi_tpcs()**. These functions make some assumptions about the start values and are not as customisable as each individual function, so we still recommend writing your own functions, but these are super useful if you want a quick idea of a few curves. Hats off to Francis Windram for the amazing additions. We will use them here to demonstrate how to parallelise fitting many models to many curves.

We will fit 5 models to all 60 TPCs in `chlorella_tpc`, demonstrating how **quickfit_multi_tpcs()** can be used with and **purrr::map()**, and how curve fitting can be sped up by parallelising using **mirai**. To switch things up, the models we will fit are _briereextended_2021_, _gaussianmodified_2006_, _briere1simplified_1999_, _tomlinsonphillips_2015_, and _lactin2_1995_.

# quickfit_tpc_multi() without parallelisation

Firstly without parallelisation. We will time how long each method takes. In **quickfit_tpc_multi()**, you can pass model specific arguments by using a list. For example, here we pass different numbers of iterations for each model, and mix a shotgun approach with a gridstart approach between models. This could be useful if you know some models are harder to fit than others.

```{r fit_many, warning=FALSE, message=FALSE}
# start timer
no_parallel_start <- Sys.time()

# fit multiple models to multiple curves
d_mods_nopara <- d %>%
  tidyr::nest(data = c(temp, rate)) %>%
  dplyr::mutate(
    mods = purrr::map(
      data,
      ~ quickfit_tpc_multi(
        data = .x,
        c(
          "briereextended_2021",
          "gaussianmodified_2006",
          "briere1simplified_1999",
          "tomlinsonphillips_2015",
          'lactin2_1995'
        ),
        "temp",
        "rate",
        iter = list(150, 150, c(4, 4, 4), 150, 200),
        lhstype = 'random'
      ),
      .progress = TRUE
    )
  ) %>%
  tidyr::unnest(mods)

# end timer
no_parallel_end <- Sys.time()
```

```{r, echo=FALSE}
cli::cli_text(cli::col_green(strrep("█", 30)), " 100% | ETA:  0s")
```
 

# quickfit_tpc_multi() with parallelisation

And now with parallelisation, which you can read more about in the [**purrr** documentation](https://purrr.tidyverse.org/reference/in_parallel.html). How and where parallelisation occurs is determined by **mirai::daemons()** that sets up daemons (persistent background processes that receive parallel computations) on your local machine or across the network.

```{r fit_many_parallel, warning=FALSE, message=FALSE}
# start timer
parallel_start <- Sys.time()

# start parallel session
daemons(3)

# fit models as before, but in parallel
d_mods_para <- d %>%
  nest(data = c(temp, rate)) %>%
  mutate(
    mods = map(
      data,
      in_parallel(
        \(x) {
          rTPC::quickfit_tpc_multi(
            data = x,
            c(
              "briereextended_2021",
              "gaussianmodified_2006",
              "briere1simplified_1999",
              "tomlinsonphillips_2015",
              'lactin2_1995'
            ),
            "temp",
            "rate",
            iter = list(150, 150, c(4, 4, 4), 150, 200),
            lhstype = 'random'
          )
        }
      ),
      .progress = TRUE
    )
  ) %>%
  tidyr::unnest(mods)

# close parallel session
daemons(0)

# end time of parallel
parallel_end <- Sys.time()

```

```{r, echo=FALSE}
cli::cli_text(cli::col_green(strrep("█", 30)), " 100% | ETA:  0s")
```

The non parallel code took `r as.numeric(no_parallel_end - no_parallel_start, units = 'secs') %>% round(0)` seconds, whereas the parallel code took `r as.numeric(parallel_end - parallel_start, units = 'secs') %>% round(0)` seconds, which is `r round((as.numeric(no_parallel_end - no_parallel_start, units = 'secs') %>% round(0))/(as.numeric(parallel_end - parallel_start, units = 'secs') %>% round(0)), 1)` times faster. That is a serious speed up!

The model fitting is likely to be the most time consuming part of the pipeline, so it is unlikely you will need to parallelise any of the other parts.

We can have a look at both objects to see that they are the same.

```{r compare}
# look at colnames
colnames(d_mods_nopara)
colnames(d_mods_para)

# look at first six rows
head(d_mods_nopara)
head(d_mods_para)
```

# Custom rTPC parallelisation

As we said, **quickfit_tpc_multi()** makes some assumptions about start values and the fitting process, so we still recommend writing your own functions. 

Here is an example of how to do this using **nls_multstart()** and **purrr** with parallelisation. We will fit **sharpeschoolhigh_1981()** and **rezende_2019()**. The overall structure is similar to we have used in `vignette("fit_many_models")` and `vignette("fit_many_curves")`.

The key difference when making things parallel is you need to specify where every function inside your custom function come from using **::** (for example **nls.multstart::nls_multstart()**.

```{r custom_fit_parallel, message=FALSE}
# write a function to fit the five models and output a tibble
fit_TPCs <- function(d, ...) {
  rezende <- nls.multstart::nls_multstart(
    rate ~ rTPC::rezende_2019(temp = temp, q10, a, b, c),
    data = d,
    iter = c(4, 4, 4, 4),
    start_lower = rTPC::get_start_vals(
      d$temp,
      d$rate,
      model_name = 'rezende_2019'
    ) -
      10,
    start_upper = rTPC::get_start_vals(
      d$temp,
      d$rate,
      model_name = 'rezende_2019'
    ) +
      10,
    lower = rTPC::get_lower_lims(d$temp, d$rate, model_name = 'rezende_2019'),
    upper = rTPC::get_upper_lims(d$temp, d$rate, model_name = 'rezende_2019'),
    supp_errors = 'Y',
    convergence_count = FALSE,
    ...
  )
  sharpeschoolhigh <- nls.multstart::nls_multstart(
    rate ~
      rTPC::sharpeschoolhigh_1981(temp = temp, r_tref, e, eh, th, tref = 15),
    data = d,
    iter = c(4, 4, 4, 4),
    start_lower = rTPC::get_start_vals(
      d$temp,
      d$rate,
      model_name = 'sharpeschoolhigh_1981'
    ) -
      10,
    start_upper = rTPC::get_start_vals(
      d$temp,
      d$rate,
      model_name = 'sharpeschoolhigh_1981'
    ) +
      10,
    lower = rTPC::get_lower_lims(
      d$temp,
      d$rate,
      model_name = 'sharpeschoolhigh_1981'
    ),
    upper = rTPC::get_upper_lims(
      d$temp,
      d$rate,
      model_name = 'sharpeschoolhigh_1981'
    ),
    supp_errors = 'Y',
    convergence_count = FALSE,
    ...
  )

  return(tibble::tibble(
    rezende = list(rezende),
    sharpeschoolhigh = list(sharpeschoolhigh)
  ))
}

# start timer
custom_parallel_start <- Sys.time()

# start parallel session
daemons(3)

# fit models
d_fits <- d %>%
  nest(data = c(temp, rate)) %>%
  mutate(
    mods = map(
      data,
      in_parallel(
        \(x) {
          fit_TPCs(d = x)
        },
        fit_TPCs = fit_TPCs
      ),
      .progress = TRUE
    )
  ) %>%
  tidyr::unnest(mods)

# close parallel session
daemons(0)

# end time of parallel
custom_parallel_end <- Sys.time()

```

```{r, echo=FALSE}
cli::cli_text(cli::col_green(strrep("█", 30)), " 100% | ETA:  0s")
```

This code took `r round(as.numeric(custom_parallel_end - custom_parallel_start, units = 'secs'), 2)` seconds to run. We can then plot all the curves as done in previous vignettes.

```{r predictions}
# create new list column of for high resolution data
d_preds <- mutate(
  d_fits,
  new_data = map(
    data,
    ~ tibble(temp = seq(min(.x$temp), max(.x$temp), length.out = 100))
  )
) %>%
  # get rid of original data column
  select(., -data) %>%
  # stack models into a single column, with an id column for model_name
  pivot_longer(
    .,
    names_to = 'model_name',
    values_to = 'fit',
    c(rezende, sharpeschoolhigh)
  ) %>%
  # create new list column containing the predictions
  # this uses both fit and new_data list columns
  mutate(preds = map2(fit, new_data, ~ augment(.x, newdata = .y))) %>%
  # select only the columns we want to keep
  select(curve_id, growth_temp, process, flux, model_name, preds) %>%
  # unlist the preds list column
  unnest(preds)

glimpse(d_preds)
```

We can the plot the predictions of each curve using **ggplot2**.

```{r plot, fig.height = 12, fig.width=8}
# plot
ggplot(d_preds) +
  geom_line(aes(temp, .fitted, col = model_name)) +
  geom_point(aes(temp, rate), d) +
  facet_wrap(~curve_id, scales = 'free_y', ncol = 6) +
  theme_bw() +
  theme(legend.position = 'none') +
  scale_color_brewer(type = 'qual', palette = 2) +
  labs(
    x = 'Temperature (ºC)',
    y = 'Metabolic rate',
    title = 'All fitted thermal performance curves',
    subtitle = 'rezende in green; sharpeschoolhigh in orange'
  )
```