---
title: "Introduction to rTPC"
author: "Daniel Padfield"
output: rmarkdown::html_vignette
date: "`r Sys.Date()`"
description: >
  An introduction to rTPC and how it can be used to fit thermal performance curves using nls.multstart.
vignette: >
  %\VignetteIndexEntry{Intro to rTPC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = 'center'
)
```

### Things to consider

- How many data points do you have for each curve? To fit all of the 23 models in **rTPC**, there needs to be a minimum of 7 points per curve.
- If you have multiple curves, it makes sense to have your data in long format, where grouping variables, temperature and rate have their own columns.
- Are there negative rate values? If a curve that crosses the x axis, it might be beneficial to only consider models that are capable of modelling negative values.
- Are there specific parameters you are interested in? If so, it may be beneficial to only consider models that explicitly include that parameter in their formulation.

***

```{r setup, message=FALSE}
# load packages
library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)
```

__rTPC__ provides a suite of functions to help fit thermal performance curves to empirical data. After searching the literature, __rTPC__ contains 23 different model formulations previously used. These functions can be easily applied to methods in R that use non-linear least squares regression to estimate thermal performance curves.

The available model formulations can be accessed using __get_model_names()__.

```{r get_mod_names}
get_model_names()
```

They are generally named after the author of the paper (and hence the name of the model within the literature) and the year at which I found the model to be first used, separated by a __"\_"__. Some original model formulations have been altered so that all models take temperature in degrees centigrade raw rate values as input.

We can demonstrate the fitting procedure by taking a single curve from the example dataset **rTPC** - a dataset of 60 TPCs of respiration and photosynthesis of the aquatic algae, *Chlorella vulgaris*. We can plot the data using **ggplot2**

```{r first_plot, fig.width=6, fig.height = 4}
# load in data
data("chlorella_tpc")

# keep just a single curve
d <- filter(chlorella_tpc, curve_id == 1)

# show the data
ggplot(d, aes(temp, rate)) +
  geom_point() +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Respiration across temperatures')

```
For each model, **rTPC** has helper functions that can estimate sensible start values (**get_start_vals()**), lower (**get_lower_lims()**) and upper (**get_upper_lims()**) limits. To demonstrate this, we shall use the sharpe-schoolfield model for high temperature inactivation only.

```{r}
# choose model
mod = 'sharpschoolhigh_1981'

# get start vals
start_vals <- get_start_vals(d$temp, d$rate, model_name = 'sharpeschoolhigh_1981')

# get limits
low_lims <- get_lower_lims(d$temp, d$rate, model_name = 'sharpeschoolhigh_1981')
upper_lims <- get_upper_lims(d$temp, d$rate, model_name = 'sharpeschoolhigh_1981')

start_vals
low_lims
upper_lims
```

One problem with most methods of fitting models in R using non-linear least squares regression is that they are sensitive to the choice of starting parameters. This problem also occurs in previous specialist R packages that help fit thermal performance curves, such as [devRate](https://github.com/frareb/devRate/) and [temperatureresponse](https://github.com/low-decarie/temperatureresponse). These methods can fail entirely or give different parameter estimates between multiple runs of the same code.

To overcome this, we recommend using the R package [**nls.multstart**](https://github.com/padpadpadpad/nls.multstart), which uses **minpackLM::nlsLM()** but allows for multiple sets of starting parameters. It iterates through multiple starting values, attempting a fit with each set of start parameters. The best model is then picked on AIC score.

In __nls_multstart__, we will use a random-search/shotgun approach to fit the curve. Random start parameter values are picked from a uniform distribution between `start_lower` and `start_upper` for each parameter. If the best model is not improved upon (in terms of AIC score) for 100 new start parameter combinations, the function will return that model fit. This is controlled by convergence_count, if this is set to FALSE, **nls_multstart()** will try and fit all iterations.

```{r fit_model}
fit <- nls_multstart(rate~sharpeschoolhigh_1981(temp = temp, r_tref,e,eh,th, tref = 15),
                                                     data = d,
                                                     iter = 500,
                                                     start_lower = start_vals - 10,
                                                     start_upper = start_vals + 10,
                                                     lower = low_lims,
                                                     upper = upper_lims,
                                                     supp_errors = 'Y')

fit
```

To estimate additional parameters of interest, we can use **rTPC::est_params()**. This function uses high resolution predictions of the fitted model to estimate traits associated with a single thermal performance curve. The currently available  methods can be viewed by running `?est_params`. For example, we may be interested in variation in the optimum temperature, $T_{opt}$, given that we adapted algae to different temperatures.

```{r}
# estimate additional traits
est_params(fit) %>%
  # round for easy viewing
  mutate_all(round, 2)

```

Finally for this introduction, we can get predictions of our model using **broom::augment()**, which is similar to **predict()**. These are then plotted over our original data.

```{r pred_and_plot, fig.width=6, fig.height = 4}
# predict new data
new_data <- data.frame(temp = seq(min(d$temp), max(d$temp), 0.5))
preds <- augment(fit, newdata = new_data)

# plot data and model fit
ggplot(d, aes(temp, rate)) +
  geom_point() +
  geom_line(aes(temp, .fitted), preds, col = 'blue') +
  theme_bw(base_size = 12) +
  labs(x = 'Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Respiration across temperatures')

```
