---
title: "Bootstrapping with weights using rTPC"
author: "Daniel Padfield"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bootstrapping with weights using rTPC}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

#### A brief example of how bootstrapping can be implemented alongside model weighting to account for model uncertainty alongside measurement uncertainty.

***
## Things to consider 

- Think carefully about your level of replication
- Bootstrapping weighted non-linear least squares is implemented using **car::Boot()**
- See `vignette("bootstrapping_models")` for examples of bootstrapping without weights
- The outcomes of the different approaches outlined here are likely to change based on the data and the model used. Use this methods with caution and think critically about their output

***

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  #tidy.opts=list(width.cutoff=60),
  #tidy=TRUE,
  fig.align = 'center',
  warning=FALSE
)
```

```{r setup, message=FALSE}
# load packages
library(boot)
library(car) # to install development version of car install.packages("car", repos = c("http://R-Forge.R-project.org"), dep = FALSE)
library(rTPC)
library(nls.multstart)
library(broom)
library(tidyverse)
library(patchwork)

```

When doing model weighting, as implemented in `vignette(model_weighting)`, the standard deviation of each point can be used to account for measurement uncertainty and can improve the model fit and reduce biases in parameter estimates. As in `vignette(bootstrapping_models`)`, **car::Boot()** can be used to do both case and residual resampling on the weighted regression model.

In case resampling, each iteration re-samples the original data with replacement (where each rate, temperature and measurement error value are linked) and uses weighted weighted non-linear least squares regression on each bootstrapped dataset.

In residual resampling of weighted non-linear regression, the weights are used as a correction to the Pearson residuals. The whole function can be viewed by using `getAnywhere(Boot.nls)`. For large samples the residual method must be ok, but the small sample properties are less clear. Ideally one would be inclined not to use the case resampling bootstrap when possible because the theory is more certain.

We will demonstrate both approaches using the example in `vignette('model_weighting')`. This vignette uses  the example dataset contained within **rTPC** - a dataset of 60 TPCs of respiration and photosynthesis of the aquatic algae, *Chlorella vulgaris*. Instead of plotting a single curve, we average over the biological replicates, `rep`, within a growth temperature (here photosynthesis from cultures adapted to 33 ºC), to get mean rate values at each assay temperature and their standard deviation. These are then plotted using **ggplot2**. Instead of fitting **kamykowski_1985**, we will instead fit **lactin2_1995** in this example.

```{r get_data, fig.height=4, fig.width = 6}
# get curve data
data("chlorella_tpc")

d_ave <- filter(chlorella_tpc, process == 'adaptation', growth_temp == 33, flux == 'photosynthesis') %>%
  group_by(temp, flux) %>%
  summarise(., sd = sd(rate),
            ave_rate = mean(rate)) %>%
  ungroup()

d_fit <- nest(d_ave, data = c(temp, ave_rate, sd)) %>%
  mutate(weighted = map(data, ~nls_multstart(ave_rate~lactin2_1995(temp = temp, a, b, tmax, delta_t),
                        data = .x,
                        iter = c(3,3,3,3),
                        start_lower = get_start_vals(.x$temp, .x$ave_rate, model_name = 'lactin2_1995') - 10,
                        start_upper = get_start_vals(.x$temp, .x$ave_rate, model_name = 'lactin2_1995') + 10,
                        lower = get_lower_lims(.x$temp, .x$ave_rate, model_name = 'lactin2_1995'),
                        upper = get_upper_lims(.x$temp, .x$ave_rate, model_name = 'lactin2_1995'),
                        supp_errors = 'Y',
                        convergence_count = FALSE,
                        # include weights here!
                        modelweights = 1/sd)))

# get predictions using augment
newdata <- tibble(temp = seq(min(d_ave$temp), max(d_ave$temp), length.out = 100))
d_preds <- d_fit %>%
  mutate(., preds = map(weighted, augment, newdata = newdata)) %>%
  select(-weighted) %>%
  unnest(preds)

# plot
ggplot() +
  geom_line(aes(temp, .fitted), d_preds) +
  geom_linerange(aes(x = temp, ymin = ave_rate - sd, ymax = ave_rate + sd), d_ave) +
  geom_point(aes(temp, ave_rate), d_ave, size = 2, shape = 21, fill = 'green4') +
  theme_bw(base_size = 12) +
  theme(legend.position = 'none',
        strip.text = element_text(hjust = 0),
        strip.background = element_blank()) +
  labs(x ='Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Photosynthesis rates across temperatures') +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  ylim(c(-0.25, 3.5))
```

***

# Method 1: Case resampling weighted non-linear regression.

In this method, each iteration re-samples the original data with replacement (where each rate, temperature and measurement error value are linked) and uses weighted weighted non-linear least squares regression on each bootstrapped dataset.

```{r case_boot}
# refit model using nlsLM
fit_nlsLM <- minpack.lm::nlsLM(ave_rate~lactin2_1995(temp = temp, a, b, tmax, delta_t),
                        data = d_ave,
                        start = coef(d_fit$weighted[[1]]),
                        lower = get_lower_lims(d_ave$temp, d_ave$ave_rate, model_name = 'lactin2_1995'),
                        upper = get_upper_lims(d_ave$temp, d_ave$ave_rate, model_name = 'lactin2_1995'),
                        weights = 1/sd)

# perform case bootstrap
boot1 <- Boot(fit_nlsLM, method = 'case')


```

We can also create model predictions for each bootstrap and confidence intervals of model predictions using the methods implemented in `vignette(bootstrapping_models)`. These can then be plotted using **ggplot2**.

```{r bootstrap1, fig.height=3,fig.width=8}
# predict over new data
boot1_preds <- boot1$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d_ave$temp), max(d_ave$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = lactin2_1995(temp, a, b, tmax, delta_t))

# calculate bootstrapped confidence intervals
boot1_conf_preds <- group_by(boot1_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975),
            .groups = 'drop')

# plot bootstrapped CIs
p1 <- ggplot() +
  geom_line(aes(temp, .fitted), d_preds, col = 'black') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot1_conf_preds, fill = 'black', alpha = 0.3) +
  geom_linerange(aes(x = temp, ymin = ave_rate - sd, ymax = ave_rate + sd), d_ave) +
  geom_point(aes(temp, ave_rate), d_ave, size = 2, shape = 21, fill = 'green4') +
  theme_bw(base_size = 10) +
  theme(legend.position = 'none',
        strip.text = element_text(hjust = 0),
        strip.background = element_blank()) +
  labs(x ='Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Photosynthesis rates across temperatures') +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  ylim(c(-3, 3.5))

# plot bootstrapped predictions
p2 <- ggplot() +
  geom_line(aes(temp, .fitted), d_preds, col = 'black') +
  geom_line(aes(temp, pred, group = iter), boot1_preds, col = 'black', alpha = 0.007) +
  geom_linerange(aes(x = temp, ymin = ave_rate - sd, ymax = ave_rate + sd), d_ave) +
  geom_point(aes(temp, ave_rate), d_ave, size = 2, shape = 21, fill = 'green4') +
  theme_bw(base_size = 10) +
  theme(legend.position = 'none',
        strip.text = element_text(hjust = 0),
        strip.background = element_blank()) +
  labs(x ='Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Photosynthesis rates across temperatures') +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  ylim(c(-3, 3.5))

p1 + p2

```

For this model formulation, the case resampling method gives really wide confidence intervals for the model predictions at high temperatures. In some of the bootstrapped datasets, the rate value at 49 ºC is not present, which results in a drastically different fit which predicts extremely negative rates for 49 ºC in those instances. As such, this is likely a bigger problem with models that allow for negative rates. Those that do not allow negative rates may not suffer so much from this issue as the mathematical formulation forces rates to always be positive, limiting the wide variation in negative rate values. However these models are also likely to have other disadvantages.

***

# Method 2: Residual resampling weighted non-linear regression.

In this method, the weights are used as a correction to the Pearson residuals of the original model fit. These modified residuals are then added onto the fitted model predictions, mimicking the methods used for weighted linear regression (Algorithm 6.3 in Davison and Hinkley (2003)). The implementation in **Boot()** can be viewed by using `getAnywhere(Boot.nls)`.

```{r bootstrap2, fig.height=3,fig.width=8}
# perform residual bootstrap
boot2 <- Boot(fit_nlsLM, method = 'residual')

# predict over new data
boot2_preds <- boot2$t %>%
  as.data.frame() %>%
  drop_na() %>%
  mutate(iter = 1:n()) %>%
  group_by_all() %>%
  do(data.frame(temp = seq(min(d_ave$temp), max(d_ave$temp), length.out = 100))) %>%
  ungroup() %>%
  mutate(pred = lactin2_1995(temp, a, b, tmax, delta_t))

# calculate bootstrapped confidence intervals
boot2_conf_preds <- group_by(boot2_preds, temp) %>%
  summarise(conf_lower = quantile(pred, 0.025),
            conf_upper = quantile(pred, 0.975),
            .groups = 'drop')

# plot bootstrapped CIs
p1 <- ggplot() +
  geom_line(aes(temp, .fitted), d_preds, col = 'black') +
  geom_ribbon(aes(temp, ymin = conf_lower, ymax = conf_upper), boot2_conf_preds, fill = 'black', alpha = 0.3) +
  geom_linerange(aes(x = temp, ymin = ave_rate - sd, ymax = ave_rate + sd), d_ave) +
  geom_point(aes(temp, ave_rate), d_ave, size = 2, shape = 21, fill = 'green4') +
  theme_bw(base_size = 10) +
  theme(legend.position = 'none',
        strip.text = element_text(hjust = 0),
        strip.background = element_blank()) +
  labs(x ='Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Photosynthesis rates across temperatures') +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  ylim(c(-3, 3.5))

# plot bootstrapped predictions
p2 <- ggplot() +
  geom_line(aes(temp, .fitted), d_preds, col = 'black') +
  geom_line(aes(temp, pred, group = iter), boot2_preds, col = 'black', alpha = 0.007) +
  geom_linerange(aes(x = temp, ymin = ave_rate - sd, ymax = ave_rate + sd), d_ave) +
  geom_point(aes(temp, ave_rate), d_ave, size = 2, shape = 21, fill = 'green4') +
  theme_bw(base_size = 10) +
  theme(legend.position = 'none',
        strip.text = element_text(hjust = 0),
        strip.background = element_blank()) +
  labs(x ='Temperature (ºC)',
       y = 'Metabolic rate',
       title = 'Photosynthesis rates across temperatures') +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  ylim(c(-3, 3.5))

p1 + p2

```

# Calculating confidence intervals of estimated and calculated parameters

As for standard non-linear regression, bootstrapping can estimate confidence intervals of the parameters explicitly modelled in the weighted regression. We can compare this approach to profiled confidence intervals (using **confint-MASS**) and asymptotic confidence intervals (using **nlstools::confint2()**).

```{r conf_int, error=TRUE}
# get parameters of fitted model
param <- broom::tidy(fit_nlsLM) %>%
  select(param = term, estimate)

# calculate confidence intervals of models
ci1 <- nlstools::confint2(fit_nlsLM, method = 'asymptotic') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'asymptotic')
ci2 <- nlstools::confint2(fit_nlsLM, method = 'profile')
# profiling method fails
ci2 <- mutate(ci1, method = 'profile',
                    conf_lower = NA,
                    conf_upper = NA)

# CIs from case resampling
ci3 <- confint(boot1, method = 'bca') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'case bootstrap')

# CIs from residual resampling
ci4 <- confint(boot2, method = 'bca') %>%
  as.data.frame() %>%
  rename(conf_lower = 1, conf_upper = 2) %>%
  rownames_to_column(., var = 'param') %>%
  mutate(method = 'residual bootstrap')

ci <- bind_rows(ci1, ci2, ci3, ci4) %>%
  full_join(., param)

ggplot(ci, aes(forcats::fct_relevel(method, c('profile', 'asymptotic')), estimate, col = method)) +
  geom_point(size = 4) +
  geom_linerange(aes(ymin = conf_lower, ymax = conf_upper)) +
  theme_bw() +
  facet_wrap(~param, scales = 'free') +
  scale_x_discrete('', labels = function(x) stringr::str_wrap(x, width = 10)) +
  labs(title = 'Calculation of confidence intervals for model parameters',
       subtitle = 'For the chlorella TPC; profile method failes')

```
Here the profiling approach fails, and the residual method returns values that are narrower than those created from asymptotic intervals. The case bootstrap does not work well. The bootstrapping approaches could also be used to estimate the uncertainty in the derived parameters, as in `vignette(bootstrapping_models)`.

#### Further reading

- John Fox (author of car) on bootstrapping regression models in R
    - https://statweb.stanford.edu/~owen/courses/305a/FoxOnBootingRegInR.pdf
- A.C. Davison & D.V. Hinkley (2003) Bootstrap Methods and their Application.         - https://www.cambridge.org/core/books/bootstrap-methods-and-their-application/ED2FD043579F27952363566DC09CBD6A
- Schenker, N., & Gentleman, J. F. (2001). On judging the significance of differences by examining the overlap between confidence intervals. The American Statistician, 55(3), 182-186.
- Puth, M. T., Neuhäuser, M., & Ruxton, G. D. (2015). On the variety of methods for calculating confidence intervals by bootstrapping. Journal of Animal Ecology, 84(4), 892-897.

